AFQ-Insight

### 2019-12-12 reports/manuscript/sections/conclusion.tex
One more pass through

We present here a novel method for analysis of dMRI tractometry data that relies
on the Sparse Group Lasso \cite{simon2013sparse} to (a) make accurate
predictions of phenotypic properties of individual subjects while,
simultaneously, (b) identifying the most important features of the white matter
in a completely data-driven approach. Importantly, through the use of a nested
cross-validation procedure, the method adapts to the constraints of small
datasets while also leveraging the power of larger datasets. Moreover, the
method is broadly applicable to a wide range of research questions: it performs
well in predicting both continuous variables, such as biological age, as well as
categorical variables, such as whether a person is a patient or a healthy control.
In both of these cases, SGL out-performs previous algorithms that have been
developed for these tasks \cite{sarica2017corticospinal, Richard2018-ux}. The
nested cross-validation approach used to fit the model and make both predictions
and inferences from it guards against overfitting and tunes the degree of
sparseness required by the algorithm. This means that both phenomena that are
locally focused in a particular anatomical location and a particular property of
diffusion (e.g., FA in the CST), as well as widely distributed phenomena can be
accurately captured.

Specifically, we demonstrated that the algorithm correctly detects the fact that
ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a targetted ROI-based approach
\cite{sarica2017corticospinal}. In contrast, in analysis of biological age, the
coefficients identified by the algorithm are very widely distributed across many
parts of the white matter, mirroring previous results with this dataset that
show a large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.

Taken together, these results demonstrate the promise of the group-regularized
regression approach. Even at the scale of dozens of subjects, the results
provided by SGL are both accurate and interpretable. Thus, this multivariate
analysis approach both (a) achieves high cross-validated accuracy for precision
medicine applications of dMRI data and (b) identifies relevant features of brain
anatomy that can further our neuroscientific understanding of clinical
disorders.

Neuroscience has entered an era in which consortium efforts are putting together
large datasets of high-quality dMRI data to address a variety of scientific
questions \cite{jernigan2016ping, jernigan2018abcd, alexander2017open,
Miller2016-hw, VanEssen2012}. The volume and complexity of these data pose a
substantial challenge. Dimensionality reduction with tractometry, followed by
analysis with the approach we present here promises to capitalize on the wealth
of data and on the co-measurement of interesting and important phenotypical data
about brain health and about the participants' cognitive abilities. We also
expect the group-regularized approach to improve with larger datasets.

@@

The results we present here also motivate extensions of the method using more
sophisticated cost functions. For example, the fused sparse group lasso (FSGL)
\cite{zhou2012} extends SGL to enforce additional spatial structure: smoothness
in the variation of diffusion metrics along the tracts. As brain measurements
include additional structure (for example, bilateral symmetry), future work
could also incorporate overlapping group membership for each entry in the tract
profiles \cite{Rao2014-xm}. For example, a measurement could come from the
corpus callosum, but also from the right hemipshere. This would also require
extending the cost function used here to incorporate these constraints.
Similarly, unsupervised dimensionality reduction \cite{Chamberland2019-mu} could
also benefit from constraints based on grouping.

>>>>

We present here a novel method for analysis of dMRI tractometry data that relies
on the Sparse Group Lasso \cite{simon2013sparse} to (a) make accurate
predictions of phenotypic properties of individual subjects while,
simultaneously, (b) identifying the features of the white matter that are most
important for this prediction in a completely data-driven approach. The method
is broadly applicable to a wide range of research questions: it performs well in
predicting both continuous variables, such as biological age, as well as
categorical variables, such as whether a person is a patient or a healthy
control. In both of these cases, SGL out-performs previous algorithms that have
been developed for these tasks \cite{sarica2017corticospinal, Richard2018-ux}.
The nested cross-validation approach used to fit the model and make both
predictions and inferences from the model guards against overfitting and tunes
the degree of sparseness required by the algorithm. This means that SGL can
accurately describe phenomena that are locally confined to a particular
anatomical location or diffusion property (e.g., FA in the CST) as well as
phenomena that are widely distributed amongst brain regions and measured
diffusion properties.

Specifically, we demonstrated that the algorithm correctly detects the fact that
ALS, which is a disease of lower motor neurons, is localized to the
cortico-spinal tract. This recapitulates the results of previous analysis of
these same data, using a targeted ROI-based approach
\cite{sarica2017corticospinal}. In contrast, in analysis of biological age, the
coefficients identified by the algorithm are very widely distributed across many
parts of the white matter, mirroring previous results with this dataset that
show a large and continuous distribution of life-span changes in white matter
properties \cite{yeatman2014lifespan}.

Taken together, these results demonstrate the promise of the group-regularized
regression approach. Even at the scale of dozens of subjects, the results
provided by SGL are both accurate, as well as interpretable
\cite{Murdoch2019-ax}: tractometry capitalizes on domain knowledge to engineer
meaningful features; SGL scores these features based on their relative
importance; enables a visualization of these feature importance scores in the
anatomical coordinate frame of the tracts (e.g., Figures \ref{fig:class-results}
and fig:regress-beta) and provides a means to understand model errors (e.g.,
Figure \ref{fig:class-errors}) . Thus, this multivariate analysis approach both
(a) achieves high cross-validated accuracy for precision medicine applications
of dMRI data and (b) identifies relevant features of brain anatomy that can
further our neuroscientific understanding of clinical disorders.

Neuroscience has entered an era in which consortium efforts are putting together
large datasets of high-quality dMRI measurments to address a variety of
scientific questions \cite{jernigan2016ping, jernigan2018abcd,
alexander2017open, Miller2016-hw, VanEssen2012}. The volume and complexity of
these data pose a substantial challenge. Dimensionality reduction with
tractometry, followed by analysis with the approach we present here promises to
capitalize on the wealth of data and on the co-measurement of interesting and
important phenotypical data about brain health and about the participants'
cognitive abilities. We also expect the group-regularized approach to improve
with larger datasets.

@@

The results we present here also motivate extensions of the method using more
sophisticated cost functions. For example, the fused sparse group lasso (FSGL)
\cite{zhou2012} extends SGL to enforce additional spatial structure: smoothness
in the variation of diffusion metrics along the tracts. As brain measurements
include additional structure (for example, bilateral symmetry), future work
could also incorporate overlapping group membership for each entry in the tract
profiles \cite{Rao2014-xm}. For example, a measurement could come from the
corpus callosum, but also from the right hemipshere. This would also require
extending the cost function used here to incorporate these constraints.
Similarly, unsupervised dimensionality reduction of tractometry data (e.g.,
\cite{Chamberland2019-mu}) could also benefit from constraints based on
grouping.
